---
title: "Data wrangling"
author: "Nicki Shumway"
date: "6/19/2019"
output: html_document
---

Loading the tidyverse
```{r}
library(tidyverse)
```

Load data
```{r}

#load the gapminder dataset

gapminder <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")

ggplot(data=gapminder, aes(y=lifeExp, x= gdpPercap)) +
  geom_point() +
  geom_hline (yintercept = median(gapminder$lifeExp), color = "red") +
  geom_vline (xintercept = median(gapminder$gdpPercap), color = "red")

# Exploring the data in R Studio - keep the data RAW!
head(gapminder)
tail(gapminder)
str(gapminder) #shows you the structure
names(gapminder) # gives you the column name

```

Look at individual variables in dataframe
```{r}

##Specify life expectancy
gapminder$lifeExp

```

Learning the filter function 'filter()'
```{r}
filter(gapminder, lifeExp < 35)

#just look at mexico
filter(gapminder, country == "Mexico", year == 2002)

#average life expectancy in Brazil between 1987 and 2007
Brazil_life_exp <-filter(gapminder, country == "Brazil", year > 1986 & year <=2007)
Brazil_life_exp
mean (Brazil_life_exp$lifeExp)

```

Learning the 'select()' function
```{r}

select(gapminder, year, country, lifeExp) ## to specifically select the columns you want
select(gapminder, -continent, -gdpPercap) ##the - indicates dropping a data column

```

Using 'filter' and 'select()' together
```{r}
gap_cambodia <- filter(gapminder, country == "Cambodia")
gap_cambodia2 <- select (gap_cambodia, -continent, -lifeExp)
```

Using the pipe ' %>% ' function (shortcut:CMD + Shift + M) %>% = and then
```{r}
gapminder %>% head()

gap_cambodia <- filter(gapminder, country == "Cambodia") %>% 
 select (-continent, -lifeExp)

## making the above even cleaner
gap_cambodia <- gapminder %>% 
  filter(country == "Cambodia") %>% #and then
 select (-continent, -lifeExp)

gap_mexico <- gapminder %>% 
  filter(country == "Mexico") %>% #and then
 select (-continent, -lifeExp)

```

Learning 'mutate()' which adds a new columb
```{r}

gapminder %>% 
  mutate(gdp = pop*gdpPercap)

##Calculate the population in thousands for all Asian countries in the year 2007 and add it as a new column

gapminder %>% 
  filter(continent == "Asia",
         year =="2007") %>% 
  mutate (pop_thousands = pop/1000)
  

```


Learning the 'group_by()' function
```{r}

gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  mutate(cont_pop = sum(pop))


```

Add in 'summarize()'
```{r}

gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  mutate(cont_pop = sum(pop)) %>% 
  summarize(cont_pop = sum(pop))

```

Add in year for each continent
```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(cont_pop = sum(pop)) %>% 
  ungroup () ## ungroup - get in the habit of ungrouping, will stay grouped for analysis
  
```

'arrange()' arranges columns 

```{r}

gapminder %>% 
  group_by(continent, year) %>% 
  summarize(cont_pop = sum(pop)) %>%
  arrange (year) %>% 
  ungroup()
```

What is the max GDP per continent across all years
```{r}

gapminder %>% 
  mutate(gdp = pop*gdpPercap) %>% 
  group_by (continent) %>%
  mutate (max_gdp = max(gdp))
  
```

