---
title: "Graphics with ggplot2"
author: "Nicki Shumway"
date: "6/18/2019"
output: html_document
---

#Loading tidyverse
```{r in tidyverse}

library(tidyverse)

```

Get the data from github

```{r}

ohi_data <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_global_data.csv")

```


#Learning ggplot2

```{r data}

ggplot(data = ohi_data, aes(x=georegion_one, y=OHI_score)) + 
  geom_jitter(width = 0.2)

```

Boxplot
```{r}

ggplot(data=ohi_data, aes (x=georegion_one)) +
  geom_bar()

```

Histograpm
```{r}
ggplot(data=ohi_data, aes (x=HDI)) +
  geom_histogram()

```

Multiple geoms exploration

```{r}

ohi_summary <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_scores_georegion_summary.csv")

ggplot(data = ohi_summary, aes(x=georegions, y=OHI_score_average)) +
  geom_bar(stat = "identity") + 
  geom_jitter(data = ohi_data, aes(x=georegion_one, y = OHI_score))

```

Adding more nuanced control over ggplot #putting geom before aes
```{r}

ggplot(data = ohi_data) +
  geom_point(aes(y = OHI_score, x = HDI, color=georegion_one)) 

```

Changing point size based on population, color based on HDI
```{r}

ggplot(data = ohi_data, aes(x = OHI_score, y = HDI, size = coastal_pop, color = HDI)) + 
  geom_point()
```

Adding a third variable (discrete)

```{r}

ggplot(data = ohi_data, aes(x = OHI_score, y = HDI, color = georegion_one)) + 
  geom_point()

```


Adding text instead of points
```{r}
ggplot(data = ohi_data, aes(x = OHI_score, y = HDI, label=country)) + 
  geom_point(aes(x = OHI_score, y = HDI)) +
  geom_text()
```


Themes
```{r}

ggplot(data = ohi_data, aes(x = OHI_score, y = HDI)) + 
  geom_point() +
  theme_bw()

```



Creating your own themes
```{r}


ggplot(data = ohi_data, aes(x = OHI_score, y = HDI)) + 
  geom_point() +
  theme_bw()

source('https://raw.githubusercontent.com/OHI-Science/ohiprep/master/src/R/scatterTheme.txt')       

ggplot(data = ohi_data, aes(x = OHI_score, y = HDI)) + 
  geom_point() + 
  scatterTheme

```

Change labels
```{r}

ggplot(data = ohi_data, aes(y = OHI_score, x = HDI, color = georegion_one)) + 
  geom_point() +
  labs(y="OHI score, 2017",
       x = "Human Development Index",
       title = "Human development is good for oceans",
       color = "Georegions")



```



Changing colors #use an established color palatte - R color Brewer

```{r}

install.packages("RColorBrewer")
library(RColorBrewer)

display.brewer.all()

my_palette <- brewer.pal(n=9, "YlOrRd") #will choose 9 colors from the YLOrRd palette


```

